---
title: 'Formula 1: Lap Times'
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
                    encoding=encoding, 
                    output_file=file.path(dirname(input_file), out_dir, 'index.html'))
  })
author: "John Goodwin"
date: "5/7/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(tidyverse)
```

```{r load_data, include=FALSE}
lap_times <- read_csv("./data/lap_times.csv")
races <- read_csv("./data/races.csv")
drivers <- read_csv("./data/drivers.csv")
results <- read_csv("./data/results.csv")

races <- races %>%
  select(raceId, year, name) %>%
  rename(race_name = name)

drivers <- drivers %>%
  mutate(driver_name = paste(forename, surname)) %>%
  select(driverId, driver_name, code) %>%
  rename(name_code = code)

## Old races allowed drivers to switch between 1 constructor's vehicles in races
## so we need to dedup to ensure join doesn't blow up
results <- results %>%
  group_by(raceId, driverId) %>%
  summarise(final_position = min(positionOrder))

lap_times <- lap_times %>%
  mutate(lap_time = milliseconds * 0.001) %>%
  left_join(races, by = "raceId") %>%
  left_join(drivers, by = "driverId") %>%
  left_join(results, by = c("raceId", "driverId")) %>%
  select(year, race_name, driver_name, name_code, lap, lap_time, final_position)

```

## Intro

This document leverages data from the [Formula 1 World Championship (1950 - 2021) Kaggle Project](https://www.kaggle.com/rohanrao/formula-1-world-championship-1950-2020) and the `plotly` R package to plot the lap times for the top 3 finishers in all of the 2020 Formula 1 races.

## R Function for Plotting

The following R function uses `plotly` to plot the lap times for the top 3 drivers. The pre-procesing of the `lap_times` data can be found in the [RMD doc](https://github.com/WiscoDisco5/f1-lap-times-plotly/blob/main/f1_lap_times.Rmd) which made this slidy presentation. The next slides use this function to plot the lap times for the 2020 F1 races.

```{r plot_function}
plot_lap_times <- function(plot_race_name) {
  lap_times %>%
    filter(year == 2020,
           race_name == plot_race_name,
           final_position <= 3)  %>%
    mutate(name_place = paste(final_position, "-", name_code)) %>%
    plot_ly(data = .,
            x = ~lap, 
            y = ~lap_time, 
            type = 'scatter', 
            mode = 'lines+markers',
            color = ~name_place) %>%
    layout(title = paste("Lap Times for Top 3 Drivers: 2020", plot_race_name),
           xaxis = list(title = "Lap Number"),
           yaxis = list(title = "Lap Time (seconds)"),
           hovermode = "x unified",
           legend = list(x = 0.1, y = 0.9))
}
```

## Austrian Grand Prix

```{r Austrian_Grand_Prix, echo=FALSE}
plot_lap_times('Austrian Grand Prix')
```

## Styrian Grand Prix

```{r Styrian_Grand_Prix, echo=FALSE}
plot_lap_times('Styrian Grand Prix')
```

## Hungarian Grand Prix

```{r Hungarian_Grand_Prix, echo=FALSE}
plot_lap_times('Hungarian Grand Prix')
```

## British Grand Prix

```{r British_Grand_Prix, echo=FALSE}
plot_lap_times('British Grand Prix')
```

## 70th Anniversary Grand Prix

```{r 70th_Anniversary_Grand_Prix, echo=FALSE}
plot_lap_times('70th Anniversary Grand Prix')
```

## Spanish Grand Prix

```{r Spanish_Grand_Prix, echo=FALSE}
plot_lap_times('Spanish Grand Prix')
```

## Belgian Grand Prix

```{r Belgian_Grand_Prix, echo=FALSE}
plot_lap_times('Belgian Grand Prix')
```

## Italian Grand Prix

```{r Italian_Grand_Prix, echo=FALSE}
plot_lap_times('Italian Grand Prix')
```

## Tuscan Grand Prix

```{r Tuscan_Grand_Prix, echo=FALSE}
plot_lap_times('Tuscan Grand Prix')
```

## Russian Grand Prix

```{r Russian_Grand_Prix, echo=FALSE}
plot_lap_times('Russian Grand Prix')
```

## Eifel Grand Prix

```{r Eifel_Grand_Prix, echo=FALSE}
plot_lap_times('Eifel Grand Prix')
```

## Portuguese Grand Prix

```{r Portuguese_Grand_Prix, echo=FALSE}
plot_lap_times('Portuguese Grand Prix')
```

## Emilia Romagna Grand Prix

```{r Emilia_Romagna_Grand_Prix, echo=FALSE}
plot_lap_times('Emilia Romagna Grand Prix')
```

## Turkish Grand Prix

```{r Turkish_Grand_Prix, echo=FALSE}
plot_lap_times('Turkish Grand Prix')
```

## Bahrain Grand Prix

```{r Bahrain_Grand_Prix, echo=FALSE}
plot_lap_times('Bahrain Grand Prix')
```

## Sakhir Grand Prix

```{r Sakhir_Grand_Prix, echo=FALSE}
plot_lap_times('Sakhir Grand Prix')
```

## Abu Dhabi Grand Prix

```{r Abu_Dhabi_Grand_Prix, echo=FALSE}
plot_lap_times('Abu Dhabi Grand Prix')
```

